\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-11-05 archive.tex
compression appendix

\documentclass{article}

\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} \usepackage{stmaryrd}% for professional tables
\usepackage{stmaryrd}(
\usepackage[section]{placeins}
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
%\usepackage{hyperref}

@@

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\denotation}[1]{{\llbracket #1 \rrbracket}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

% If accepted, instead use the following line for the camera-ready submission:
%\usepackage[accepted]{icml2018}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}

@@

\twocolumn[
\icmltitle{\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2018
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Kevin Ellis}{MIT}
\icmlauthor{Lucas Morales}{MIT}
\icmlauthor{Mathias}{MIT}
\icmlauthor{Maxwell Nye}{MIT}
\icmlauthor{Armando Solar-Lezama}{MIT}
\icmlauthor{Joshua B. Tenenbaum}{MIT}
\end{icmlauthorlist}

%\icmlaffiliation{to}{Department of Computation, University of Torontoland, Torontoland, Canada}
\icmlaffiliation{MIT}{MIT}
%\icmlaffiliation{ed}{School of Computation, University of Edenborrow, Edenborrow, United Kingdom}

\icmlcorrespondingauthor{Kevin Ellis}{ellisk@mit.edu}
%\icmlcorrespondingauthor{Eee Pppp}{ep@eden.co.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

%\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

%% \begin{abstract}
%%   Successful approaches to program induction require a hand-engineered
%%   domain-specific language (DSL), constraining the space of allowed
%%   programs and imparting prior knowledge of the domain.  We contribute
%%   a program induction algorithm called \system that learns a DSL while
%%   jointly training a neural network to efficiently search for programs
%%   in the learned DSL.\@ We use our model to synthesize functions on lists,
%%   edit text, and solve symbolic regression problems, showing how the
%%   model learns a domain-specific library of program components for
%%   expressing solutions to problems in the domain.
%% \end{abstract}

@@

The purpose of training the recognition model is to amortize the cost
of searching for programs.  It does this by learning to predict, for
each task, programs with high likelihood according to
$\probability[x|p]$ while also being probable under the prior
$(\mathcal{D},\theta)$.  Concretely, the recognition model $q$
predicts, for each task $x\in X$, a weight vector $q(x) =
\theta^{(x)}\in \mathbb{R}^{|\mathcal{D}| + 1}$.  Together with the
DSL, this defines a distribution over programs,
$\probability[p|\mathcal{D},\theta = q(x)]$.  We abbreviate this
distribution as $q(p|x)$.  The crucial aspect of this framing is that
the neural network leverages the structure of the learned DSL, so it
is \emph{not} responsible for generating programs wholesale.  We share
this aspect with DeepCoder~\cite{balog2016deepcoder}
and~\cite{menon2013machine}.

How should we get the data to train $q$?
This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered during waking.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

@@

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
Intuitively, we want the algorithm to look at the programs found
during waking and generalize beyond them, both so the DSL can better
express the current solutions, and also so that the DSL might expose
new abstractions which will later be used to discover more programs.

Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{AIC}
\end{align}
At a high level, our approach is to search locally through the space
of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
fails to increase.  These small changes consist of introducing new
candidate $\lambda$-expressions into the DSL.

@@

Formally, we want a DSL
which puts high probability on
any refactoring of the programs found during waking, maximizing:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{factorObjective}
\end{align}
where $p'\manyReduce p$
is the standard notation for ``expression $p'$ evaluates to $p$ by the standard rules of $\lambda$-calculus''~\cite{pierce}.

@@

\textbf{Idea 1:} Limit the degree to which
a piece of code can be refactored.
Instead of considering every refactoring,
bound the number of $\lambda$-calculus evaluation steps
separating a refactoring from its original program.
Formally,
we define the set of $n$-step refactorings as:
\begin{equation}
  R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
\end{equation}
where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
For example,
\begin{align*}
  \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
  \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
  \code{(lambda (y) y)}&
\end{align*}
Returning to Equation~\ref{factorObjective},
this approximation gives the following objective:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\+&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{limitedObjective}
\end{align}
In practice, setting the number of refactoring steps $n$ to 3 suffices
to give a competent DSL learning algorithm.  Although the number of
refactorings is now finite, it is still prohibitively large, and in
fact grows exponentially quickly both as a function of $n$ and as a
function of the size of the program being refactored.  Next, we show
how to tame this exponential explosion.

\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
we developed a technique for compactly representing and efficiently
calculating the set of all refactorings.  The key idea is to represent
the set of refactorings using a \emph{version
  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
A version space is a tree-shaped data structure
that compactly represents a large set of programs.
We first define version spaces over $\lambda$-calculus expressions,
and then give a dynamic program for constructing a version space
that represents the set of possible refactorings (i.e., $R_n(p)$).

A version space is either:
\begin{itemize}
\item A deBuijn index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item A  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}\\
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}\\
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}\\
    \denotation{\varnothing}& = \varnothing\\
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as $e\in v$, and which is equivalent to $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new DSL.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align*}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in v$}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}\\
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\ell(e|\mathcal{D})
  \end{align*}
where $\ell(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\ell(p|\mathcal{D})$.

@@

With this machinery in hand, we now have all the pieces needed to
learn a DSL (Algorithm~\ref{grammarInductionAlgorithm})

    \begin{algorithm*}[H]
    %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
       \caption{DSL Induction Algorithm}
       \label{grammarInductionAlgorithm}
       \begin{algorithmic}[1]
         \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
%         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
         \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
         \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
         \WHILE {true}         
         \STATE $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
         \STATE Define $\text{refactor}(p|\mathcal{D}') = \argmin_{e\in \denotation{v_p}} \ell (e|\mathcal{D}')$ \Comment{Rewrites $p$ assuming $\mathcal{D}'$ is the DSL}
         \STATE Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
         \STATE Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
         \STATE Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
         \STATE components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
         \STATE proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
         \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
         \STATE $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
         \STATE $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
         \ENDWHILE
       \end{algorithmic}
    \end{algorithm}

@@

\bibliography{main}
\bibliographystyle{icml2018}

>>>>

\documentclass{article}
\usepackage[margin=1in]{geometry}
\usepackage{dblfloatfix}    % To enable figures at the bottom of page
% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} \usepackage{stmaryrd}% for professional tables
\usepackage{stmaryrd}(
\usepackage[section]{placeins}
% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2018} with \usepackage[nohyperref]{icml2018} above.
%\usepackage{hyperref}

@@

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\denotation}[1]{{\llbracket #1 \rrbracket}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
%\usepackage[nohyperref]{icml2018}

@@

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

@@

@@

The purpose of training the recognition model is to amortize the cost
of searching for programs.  It does this by learning to predict, for
each task, programs with high likelihood according to
$\probability[x|p]$ while also being probable under the prior
$(\mathcal{D},\theta)$.  Concretely, the recognition model $Q$
predicts, for each task $x\in X$, a distribution over programs,
$Q(\cdot |x)$.
We want $Q(\cdot |x)$ to approximate the posterior $\probability[\cdot |x,\mathcal{D},\theta]$.

\subsubsection{Training $Q$}
How should we get the data to train $q$?
This is nonobvious  because we are considering a weakly supervised setting (i.e., learning only from tasks and not from (program, task) pairs).
One approach is to sample programs from the DSL,
run them to get their input/outputs,
and then train $q$ to predict the program from the input/outputs.
This is like how a Helmholtz machine
trains its recognition model during its ``sleep'' phase~\cite{dayan1995helmholtz}.
The advantage of ``Helmholtz machine'' training is that
we can draw unlimited samples from the DSL,
training on a large amount of data.
Another approach is
 self-supervised learning,
training $q$ on the (program, task)
pairs discovered during waking.
The advantage of self-supervised learning
is that the training data is much higher quality,
because we are training on the actual tasks.
Due to these complementary advantages,
we train on both these sources of data.

@@

\subsubsection{Parameterizing $Q$}

Broadly the literature contains two different approaches to
parameterizing conditional distributions over programs.  The first
approach, seen in systems like RobustFill~\cite{devlin2017robustfill}
and~\cite{zavershynskyi2018naps}, is to use a recurrent neural network
to predict the program token-by-token ala Seq2Seq or Seq2Tree.  The
advantage of this approach is that the neural net can predict the
entire program, so if the network is sufficiently powerful, it can
completely solve the synthesis problem.  There are two disadvantages
to this approach.  First, these models perform poorly at out-of-sample
generalization~\cite{}, which is critical for our setting, as the
agent may need to solve new tasks that are qualitatively different
from the tasks it has solved so far.  Second, a powerful deep
recurrent network may be costly to sample or enumerate from ---
so if the network cannot easily solve a task, we cannot compensate
with rapid sampling or enumeration.  In contrast, state-of-the-art
enumerative program synthesizers evaluate millions of programs per
second~\cite{feser2015synthesizing}.

The second approach is to have $Q$ predict a fixed-dimensional weight
vector, which then biases a fast enumerator~\cite{balog2016deepcoder}
or sampler~\cite{menon2013machine}.  This approach can enjoy strong
out-of-sample generalization, because it can fall back on enumeration
or sampling when the target program is unlike the training programs.
A main drawback is that the neural net is deliberately handicapped,
and can only send so much information about the target program.

We adopt a middle ground between these two extremes.  Our recognition
model predicts a distribution over primitives in the DSL,
conditioned on the local context in the syntax tree of the
program. When predicting the next node to add to the syntax tree of a program,
the recognition model
conditions on the parent node, as well as
which argument is being generated.
This is a kind of `bigram' model over trees,
where the bigrams
take the form of (parent, child, argument index).
See Figure~\ref{bg}.
This parameterization
has three main advantages:
(1) it supports fast enumeration
and sampling of programs,
because the
recognition model
only needs to run once for each task;
(2) it allows the recognition model to provide fine-grained
information about the structure of the target program;
and (3)
training this recognition model
causes it to learn to break symmetries in the space of programs,
described next.
\begin{figure}
\centering  \begin{tikzpicture}%[every node/.style={inner sep=0,outer sep=0}]
    \node(l)[draw] at (0,0) {\code{$\lambda$ (x)}};
    \node(k)[draw] at ([yshift=-50]l.south) {\code{+}};
    \node(o)[draw] at ([xshift=-50,yshift=-50]k.south) {\code{1}};
    \node(m)[draw] at ([xshift=50,yshift=-50]k.south) {\code{*}};
    \node(x1)[draw] at ([xshift=50,yshift=-50]m.south) {\code{x}};
    \node(x2)[draw] at ([xshift=-50,yshift=-50]m.south) {\code{x}};

    \draw[->] (l.south)-- node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{+}|\text{start})$}(k.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{1}|\code{+}_1)$}(o.north);
    \draw[->] (k.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{*}|\code{+}_2)$}(m.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{x}|\code{*}_2)$}(x1.north);
    \draw[->] (m.south)--node[fill=white,align=center,midway,inner sep=0,outer sep=0]{$q(\code{x}|\code{*}_1)$}(x2.north);
\end{tikzpicture}
\caption{Parameterization of distribution over programs predicted by recognition model.
  Here the program (syntax tree shown above) is \code{($\lambda$ (x) (+ 1 (* x x)))}.
Each conditional distribution predicted by the recognition model is written $q(\text{child}|\text{parent}_{\text{argument index}})$.}
  \end{figure}

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
Intuitively, we want the algorithm to look at the programs found
during waking and generalize beyond them, both so the DSL can better
express the current solutions, and also so that the DSL might expose
new abstractions which will later be used to discover more programs.
Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{AIC}
\end{align}
At a high level, our approach is to search locally through the space
of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
fails to increase.  These small changes consist of introducing new
candidate $\lambda$-expressions into the DSL.

@@

Formally, we want a DSL
which puts high probability on
any refactoring of the programs found during waking, maximizing:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{factorObjective}
\end{align}
where $p'\manyReduce p$
is the standard notation for ``expression $p'$ evaluates to $p$ by the rules of $\lambda$-calculus''~\cite{pierce}.

@@

\textbf{Idea 1:} Limit the degree to which
a piece of code can be refactored.
Instead of considering every refactoring,
bound the number of $\lambda$-calculus evaluation steps
separating a refactoring from its original program.
Formally,
we define the set of $n$-step refactorings as:
\begin{equation}
  R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
\end{equation}
where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
For example,
\begin{align*}
  \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
  \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
  \code{(lambda (y) y)}&
\end{align*}
Returning to Equation~\ref{factorObjective},
this approximation gives the following objective:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\+&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{limitedObjective}
\end{align}
In practice, setting the number of refactoring steps $n$ to 3 suffices
to give a competent DSL learning algorithm.  Although the number of
refactorings is now finite, it is still prohibitively large, and in
fact grows exponentially quickly both as a function of $n$ and as a
function of the size of the program being refactored.  For example,
for the programs in Figure~\ref{mapFactor}, there are approximately
$10^{14}$ possible refactorings.  Next, we show how to tame this
exponential explosion.

\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
we developed a technique for compactly representing and efficiently
calculating the set of all refactorings.  The key idea is to represent
the set of refactorings using a \emph{version
  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
A version space is a tree-shaped data structure
that compactly represents a large set of programs.
We first define version spaces over $\lambda$-calculus expressions,
and then give a dynamic program for constructing a version space
that represents the set of possible refactorings (i.e., $R_n(p)$).
This technique is astronomically more efficient than
explicitly representing
the space of possible refactorings:
for the example in Figure~\ref{mapFactor},
we represent
the space of refactorings using a version space with
$10^6$ nodes,
which encodes $10^{14}$ refactorings.
Formally, a version space is either:
\begin{itemize}
\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus. When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to. For example, $\lambda x.\lambda y. (x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices. See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}\\
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}\\
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}\\
    \denotation{\varnothing}& = \varnothing\\
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as $e\in v$, and which is equivalent to $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new DSL.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in v$. Exploits the fact that $e\in v$ can be efficiently computed.}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}\label{FACTORING}\\\nonumber
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\\nonumber
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\\nonumber
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
  \end{align*}
where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.

@@

With this machinery in hand, we now have all the pieces needed to
learn a DSL (Algorithm~\ref{grammarInductionAlgorithm}).
To define the prior distribution over $(\mathcal{D},\theta})$ (Algorithm~\ref{grammarInductionAlgorithm}, lines 7-8), we penalize the syntactic complexity of the $\lambda$-calculus expressions in the DSL, defining $    \probability[\mathcal{D}]\propto\exp(-\lambda\sum_{p\in \mathcal{D}}\text{size}(p) )$ where $\text{size}(p)$  measures the size of the syntax tree of program $p$,
  and $\lambda$ controls how strongly we regularize the size of the DSL.
  We place a symmetric Dirichlet prior over the weight vector $\theta$.

To appropriately score each proposed $\mathcal{D}$ we must reestimate
 the weight vector $\theta$ (Algorithm~\ref{grammarInductionAlgorithm}, line 7).
Although this  may seem 
very similar to estimating the parameters of a probabilistic context free grammar,
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
our DSLs are context-sensitive due to the presence of variables
in the programs and also due to the polymorphic typing system.
Appendix~\ref{mapAppendix} derives a tractable MAP estimator for $\theta$.

\begin{algorithm}%[H]
  %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
  \caption{DSL Induction Algorithm}
  \label{grammarInductionAlgorithm}
  \begin{algorithmic}[1]
    \State {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
    %         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
    \State \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
    \State $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
    \While{true}         
    \State $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
    \State Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
    \State Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
    \State Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
    \State components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
    \State proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
    \State $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
    \State \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
    \State $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
    \State $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
    \EndWhile
  \end{algorithmic}
\end{algorithm}

@@

\bibliography{main}
\bibliographystyle{plain}
